---
title: "Visualizations LDA coefficients"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library("ggpubr")
library(LDATS)
library(ggVennDiagram)
library(stringr)
library(abind)
library(patchwork)

source("utils.R")
```
Visualizations on three sets of networks. First set consists of networks trained in 30 replications on CIFAR 10 training set with 500 sample validation set extracted randomly in each replication. Second set consists of networks trained in a single replication on half of CIFAR 10 training set. Third set are networks trained in 10 replications on half of CIFAR 100 training set.


## CIFAR 10

```{r}
base_dir <- "../data/data_train_val_c10"
repls <- 0:29
classes <- 10
lda_coefs <- load_lda_coefs(base_dir, repls)
net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
ens_pw_cal <- read.csv(file.path(base_dir, "ensemble_pw_calibration.csv"))
ens_pw_irrel <- read.csv(file.path(base_dir, "ensemble_pw_irrelevant.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
ens_pw_cal$bin_c <- (ens_pw_cal$conf_min + ens_pw_cal$conf_max) / 2
ens_pw_cal[, c("class1", "class2", "bin_c")] <- lapply(ens_pw_cal[, c("class1", "class2", "bin_c")], as.factor)
ens_pw_irrel[, c("class1", "class2")] <- lapply(ens_pw_irrel[, c("class1", "class2")], as.factor)

```

```{r, fig.width=10, fig.height=8}
for (cl1 in 1:(classes - 1))
{
  for (cl2 in (cl1 + 1):classes)
  {
    lda_plt <- lda_coefs %>% filter(class1 == cl1 & class2 == cl2) %>% ggplot() + geom_boxplot(aes(x=coefficient, y=value)) +
      facet_wrap(~train_type) + ggtitle("LDA coefficients")
    
    acc_plt_net <- net_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(x=network, y=accuracy)) + 
      geom_boxplot() + ggtitle("Pairwise accuracies of networks")
    
    acc_plt_ens <- ens_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(y=accuracy)) +
      geom_boxplot() + facet_wrap(~train_set) + ggtitle("Pairwise accuracies of LDA output") +
      theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
    
    
    print((acc_plt_net/lda_plt/acc_plt_ens) + plot_annotation(title=paste("Classes ", cl1, " vs ", cl2)))
  }
}
```

An attempt to visualize calibration of LDAs. Almost all elements are present near the zero and one probabilities, therefore the visualization is not very informative.
```{r}
for (cl1 in 1:(classes - 1))
{
  for (cl2 in (cl1 + 1):classes)
  {
    cal_plt <- ens_pw_cal %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot() +
      geom_boxplot(mapping=aes(x=bin_c, y=bin_accuracy)) + facet_wrap(~train_set) +
      ggtitle(paste("Classes ", cl1, " vs ", cl2)) + xlab("Confidence") + ylab("Accuracy")
    
    print(cal_plt)
  }
}
```
```{r}
for (cl1 in 1:(classes - 1))
{
  for (cl2 in (cl1 + 1):classes)
  {
    irrel_plt <- ens_pw_irrel %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot() +
      geom_histogram(mapping=aes(x=pred1), bins=30) + facet_wrap(~train_set) +
      ggtitle(paste("Predictions for irrelevant classes by LDA ", cl1, " vs ", cl2)) +
      xlab("Predicted probability")
    
    print(irrel_plt)
  }
}
```
These histograms show, that even for unknown classes, LDA most osten produces probabilities close to one and zero. One exception is val_train LDA for class pair 4 and 6, where probabilities are divided evenly. Also the confidence plot for this LDA looks sensibly.


## CIFAR 10 half

```{r}
base_dir <- "../data/data_train_val_half_c10"
repls <- 0:0
folds <- 0:49
classes <- 10
lda_coefs <- load_lda_coefs(base_dir, repls, folds)
net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
```

```{r, fig.width=10, fig.height=8}
for (cl1 in 1:(classes - 1))
{
  for (cl2 in (cl1 + 1):classes)
  {
    lda_plt <- lda_coefs %>% filter(class1 == cl1 & class2 == cl2) %>% ggplot() + geom_boxplot(aes(x=coefficient, y=value)) +
      facet_wrap(~train_type) + ggtitle("LDA coefficients")
    acc_plt_net <- net_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(x=network, y=accuracy)) + 
      geom_boxplot() + ggtitle("Pairwise accuracies of networks")
    acc_plt_ens <- ens_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(y=accuracy)) +
      geom_boxplot() + facet_wrap(~train_set) + ggtitle("Pairwise accuracies of LDA output") +
      theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
    
    
    print((acc_plt_net/lda_plt/acc_plt_ens) + plot_annotation(title=paste("Classes ", cl1, " vs ", cl2)))
  }
}
```

## CIFAR 100 half

```{r}
base_dir <- "../data/data_train_val_half_c100"
repls <- 0:9
folds <- 0:4
classes <- 10
lda_coefs <- load_lda_coefs(base_dir, repls, folds)
net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
```

```{r, fig.width=10, fig.height=8}
for (cl1 in 1:(classes - 1))
{
  for (cl2 in (cl1 + 1):classes)
  {
    lda_plt <- lda_coefs %>% filter(class1 == cl1 & class2 == cl2) %>% ggplot() + geom_boxplot(aes(x=coefficient, y=value)) +
      facet_wrap(~train_type) + ggtitle("LDA coefficients")
    acc_plt_net <- net_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(x=network, y=accuracy)) + 
      geom_boxplot() + ggtitle("Pairwise accuracies of networks")
    acc_plt_ens <- ens_pw_results %>% filter(class1 == (cl1 - 1) & class2 == (cl2 - 1)) %>% ggplot(mapping=aes(y=accuracy)) +
      geom_boxplot() + facet_wrap(~train_set) + ggtitle("Pairwise accuracies of LDA output") +
      theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
    
    
    print((acc_plt_net/lda_plt/acc_plt_ens) + plot_annotation(title=paste("Classes ", cl1, " vs ", cl2)))
  }
}
```

Here are too many class pair combinations to draw them all. TODO: think of some criteria to draw interesting ones. 
