---
title: "Outputs inspection half CIFAR100"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library("ggpubr")
library(LDATS)
library(ggVennDiagram)
library(stringr)
library(abind)
library(patchwork)

source("utils.R")
```

Visualization on both CIFAR 10 and 100.
To better understand the differences between train_training and val_training LDA training methodologies, we focus on pairwise predictions. This allows us to observe the outputs of LDA models without interference from the coupling methods.

## CIFAR 10 - validation set of size 500

```{r}
base_dir <- "../data/data_train_val_c10"
repls <- 0:29
folds <- 0:0
classes <- 10

net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
```

```{r}
for (ri in repls)
{
  net_plot <- net_pw_results %>% filter(repli == ri) %>% 
    ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~network) + 
    xlab("Class") + 
    ylab("Class") +
    scale_y_discrete(limits=rev) +
    scale_fill_binned(type="viridis", limits=c(0.92, 1),  name="accuracy") +
    coord_fixed() + 
    ggtitle("Pairwise accuracies networks") +
    theme(plot.title = element_text(hjust = 0.5),
      axis.ticks = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())
  
  ens_plot <- ens_pw_results %>% filter(repli == ri & fold == 1) %>% 
    ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~train_set) + 
    xlab("Class") + 
    ylab("Class") +
    scale_y_discrete(limits=rev) +
    coord_fixed() +
    ggtitle("Pairwise accuracies ensembles") +
    scale_fill_binned(type="viridis", limits=c(0.92, 1), name="accuracy") +
    theme(plot.title = element_text(hjust = 0.5),
      axis.ticks = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())
    
    print((net_plot/ens_plot) + plot_annotation(title=paste("Replication ", ri)))
}
```
LDA combination is able to improve pairwise probabilities in almost all examples. tt seems to be a little bit more successful than vt. Especially problematic seem classes 3 and 5 - cat and dog.

## CIFAR 10 - validation set of size 25000

```{r}
base_dir <- "../data/data_train_val_half_c10"
repls <- 0:0
folds <- 0:49
classes <- 10

net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
```

```{r}
for (ri in repls)
{
  net_plot <- net_pw_results %>% filter(repli == ri) %>% 
    ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~network) + 
    xlab("Class") + 
    ylab("Class") +
    scale_y_discrete(limits=rev) +
    scale_fill_binned(type="viridis", limits=c(0.92, 1),  name="accuracy") +
    coord_fixed() + 
    ggtitle("Pairwise accuracies networks") +
    theme(plot.title = element_text(hjust = 0.5),
      axis.ticks = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())
  
  
  
  for (fi in folds)
  {
    ens_plot <- ens_pw_results %>% filter(repli == ri & fold == fi) %>% 
      ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~train_set) + 
      xlab("Class") + 
      ylab("Class") +
      scale_y_discrete(limits=rev) +
      coord_fixed() +
      ggtitle("Pairwise accuracies ensembles") +
      scale_fill_binned(type="viridis", limits=c(0.92, 1), name="accuracy") +
      theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
    
    print((net_plot/ens_plot) + plot_annotation(title=paste("Replication ", ri, ", fold ", fi)))
  }
}
```
Networks are same across all the folds. vt and tt for a given fold have no special relationship, they are displayed together just for convenience.
tt again seems to be more successfull at improving the parwise probabilities.

## CIFAR 10 - validation set of size 25000

```{r}
base_dir <- "../data/data_train_val_half_c100"
repls <- 0:9
folds <- 0:4
classes <- 100

net_pw_results <- read.csv(file.path(base_dir, "net_pw_accuracies.csv"))
ens_pw_results <- read.csv(file.path(base_dir, "ensemble_pw_accuracies.csv"))
net_pw_results[, c("class1", "class2")] <- lapply(net_pw_results[, c("class1", "class2")], as.factor)
ens_pw_results[, c("class1", "class2")] <- lapply(ens_pw_results[, c("class1", "class2")], as.factor)
```

```{r}
for (ri in repls)
{
  net_plot <- net_pw_results %>% filter(repli == ri) %>% 
    ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~network) + 
    xlab("Class") + 
    ylab("Class") +
    scale_y_discrete(limits=rev, breaks=seq(0, classes, 20)) +
    scale_x_discrete(breaks=seq(0, classes, 20)) +
    scale_fill_binned(type="viridis",  name="accuracy", limits=c(0.7, 1)) +
    coord_fixed() + 
    ggtitle("Pairwise accuracies networks") +
    theme(plot.title = element_text(hjust = 0.5),
      axis.ticks = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())
  
  
  
  for (fi in folds)
  {
    ens_plot <- ens_pw_results %>% filter(repli == ri & fold == fi) %>% 
      ggplot(mapping=aes(x=class2, y=class1, fill=accuracy)) + geom_raster() + facet_wrap(~train_set) + 
      xlab("Class") + 
      ylab("Class") +
      scale_y_discrete(limits=rev, breaks=seq(0, classes, 20)) +
      scale_x_discrete(breaks=seq(0, classes, 20)) +
      coord_fixed() +
      ggtitle("Pairwise accuracies ensembles") +
      scale_fill_binned(type="viridis", name="accuracy", limits=c(0.7, 1)) +
      theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
    
    print((net_plot/ens_plot) + plot_annotation(title=paste("Replication ", ri, ", fold ", fi)))
  }
}
```

Need some other way to visualize this.